### 2023-07-05 1.Introduction/1.motivation.tex
Bunch of changes, mainly great advance in lstm state of the art

Another use for \ac{AI} that works great in embedded use is for camera relocalization \cite{videoReloc}. That is useful especially for VR headsets that need to know their location in the room. Having an embedded chip to do all this computation.

>>>>

Another use for \ac{AI} that works great in embedded use is for camera relocalization \cite{videoReloc}. That is useful especially for VR headsets that need to know their location in the room. The embedded chip would allow this computation to be much faster than other methods, and thus give more realistic results.

The point of this thesis is to lay out the ground work to create a LSTM capable chip to use in embedded systems. It could be used detection and prediction of any kind of sequential events.

### 2023-07-05 2.State/lstm.tex
Working on lstm
Bunch of changes, mainly great advance in lstm state of the art

\acp{LSTM} are able to forget information passed. This ability gave the uncommon name of \acl{LSTM} as it has both long and short term memory. This what gives \acp{LSTM} their use for sequence data. They can analyze the data and keep the information from the last time step to make a better decision afterwards. The most comprehensible example is considering a sentence. (TODO : find example)

@@

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{lstmCellLatex.svg}
  \caption{LSTM Cell}
  \label{fig:lstmCell}
\end{figure}

Figure \ref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture. In an \ac{LSTM}, each gate is a different \ac{NN}.

An \ac{LSTM} works using different gates

>>>>

\acp{LSTM} were born from the vanishing/exploding gradient problem. Traditionnal \ac{RNN} have the ability to model sequential events by propagating through time, for example forward and backward propagation. This is achieved by connecting these sequential events with the hidden state:
$$h_t=f(h_{t-1},x_t)$$
The hidden state $h_t$ carries all the past informations for the next time step.
Even though \acp{RNN} work very well, they are subject to the vanishing/exploding gradient problem. This issue occurs when too many time steps are chained.

\acp{LSTM} alleviate this issue by adding a cell state, this state gives it the ability to choose what input is important and which one isn't, thus giving it a long term memory. This ability gave the uncommon name of \acl{LSTM} as it has both long and short term memory. This what gives \acp{LSTM} their use for sequence data. They can analyze the data and keep the information from the last time step to make a better decision afterwards. The most comprehensible example is considering a sentence. (TODO : find example)

@@

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{lstm/lstmCell.svg}
  \caption{LSTM Cell}
  \label{fig:lstmCell}
\end{figure}

Figure \ref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture. In an \ac{LSTM}, each gate is a different \ac{NN} and then activated with either a tanh or a sigmoid activation function. Each input to the cell is a vector.
Those vector are of varying sizes depending on several factors. For example, both $h_t$ and $c_t$ are of the same size as the number of hidden state (sometimes refered to as cell state) for any $t\geq 0$.
The input vector ($x_t$) is of size of the sample we want to feed for each time step.

\subsection{Equations}

The equations of an LSTM are quite unusual.
Let's start with the more classic gate equations. They are the ones that behave like the more classic \ac{NN}.
The input (equation \ref{eq:inputG}), forget (equation \ref{eq:forgetG}) and ouput (equation \ref{eq:ouputG}) gates are described below.

\begin{equation}\label{eq:inputG}
  i_t=\sigma (w_i\cdot[x_{t_1},h_{t-1}] + b_i)
\end{equation}
\begin{equation}\label{eq:forgetG}
  f_t=\sigma (w_f\cdot[x_{t_1},h_{t-1}] + b_f)
\end{equation}
\begin{equation}\label{eq:ouputG}
  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}] + b_o)
\end{equation}

Where ($w_i$,$b_i$), ($w_f$,$b_f$) and ($w_o$,$b_o$) are the pair of weights and biases for the input, forget and ouput gates respectively. $x_t$ is the input vector and $h_t$ is the hidden state vector.

For the next equation describes the candidate cell state (equation \ref{eq:candCell}), that will next become the cell state (equation \ref{eq:cellS}).

\begin{equation}\label{eq:candCell}
  \tilde{c}_t=tanh(w_c\cdot[x_{t-1},h_{t-1}] + b_c)
\end{equation}
\begin{equation}\label{eq:cellS}
  c_t=f_t\cdot c_{t-1} + i_t \cdot \tilde{c}_t
\end{equation}

Where $w_c$ and $b_c$ are the weights and bias for the candidate cell state.

The final step of the \ac{LSTM} is to compute the hidden state (equation \ref{eq:hiddenS}).
\begin{equation}\label{eq:hiddenS}
  h_t=o_t\cdot tanh(c_t)
\end{equation}

We set $x_0$ as the first input and define $h_0$ as a zero only vector.

### 2023-07-05 2.State/nn.tex
Working on lstm
Bunch of changes, mainly great advance in lstm state of the art

% For the State of the art

\ac{AI} is wide topic, usually when we talk about \acp{AI}, we refer to the most common type, Machine Learning. Especially, a sub part of Machine Learning called Deep Learning. Deep Learning has seen a surge in research the last decade, which lead to great advancement in this field and the numerous \ac{AI} tools that popped up in the last years.
Deep Learning works using what is called \acp{NN}.

Those \acp{NN} require matrix multiplication. This kind of operation is very time and energy consuming to run on a classic computer. There are options to better run those operations. GPUs can be better suited than a CPU for a \ac{NN} application as it is made for 3D graphics that also mainly use Matrix Multiplication. Other specialized hardware such as FPGAs and ASICs can improve both execution time and energy consumption. All those are digital computers, using analog computing could highly improve speed, power consumption and even smaller chip area. All this using a memristive crossbar array \cite{Xbar}. An analog computer
The point of this thesis is to lay out the ground work to create a LSTM capable chip to use in embedded systems. It could be used detection and prediction of any kind of sequential events.
Using those memristors \cite{TheoMemristor}, which work like resistance with a memory component to them, that means that they can be set to a desired value.

@@

@@

In a computational \ac{NN}, synapses are represented by weights. Those weights are to be multiplied by the last neuron and then added to each other to produce the next stage. Using the names defined in figure \ref{fig:snn}, the equation linking each layer is the following matrix multiplication :
\begin{equation}
  \begin{bmatrix}
    H1\\ H2\\ H3\\ H4\\
  \end{bmatrix}
  =
  \begin{bmatrix}
    W1,1 & W1,2 & W1,3\\
    W2,1 & W2,2 & W2,3\\
    W3,1 & W3,2 & W3,3\\
    W4,1 & W4,2 & W4,3\\
  \end{bmatrix}
  \cdot
  \begin{bmatrix}
    I1\\ I2\\ I3\\
  \end{bmatrix}
\end{equation}

>>>>

@@

\ac{AI} is wide topic, it refers to any computer system that can make a descision. It can take many forms but usually when we talk about \acp{AI}, we refer to the most common type, Machine Learning. Especially, a sub part of Machine Learning called Deep Learning. Deep Learning has seen a surge in research the last decade, which lead to great advancement in this field and the numerous \ac{AI} tools that popped up in the last years.
Deep Learning works using what is called \acp{NN}.

@@

In a computational \ac{NN}, synapses are represented by weights. Those weights are to be multiplied by the last neuron and then added to each other to produce the next stage. Using the names defined in figure \ref{fig:snn}, the equation linking each layer is the following matrix multiplication :
\begin{equation}
  \begin{bmatrix}
    H_1\\ H_2\\ H_3\\ H_4\\
  \end{bmatrix}
  =
  \begin{bmatrix}
    W_{1,1} & W_{1,2} & W_{1,3}\\
    W_{2,1} & W_{2,2} & W_{2,3}\\
    W_{3,1} & W_{3,2} & W_{3,3}\\
    W_{4,1} & W_{4,2} & W_{4,3}\\
  \end{bmatrix}
  \cdot
  \begin{bmatrix}
    I_1\\ I_2\\ I_3\\
  \end{bmatrix}
\end{equation}
